A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-11 tex/codatadp.tex
more

Finally, some codatatypes are so degenerate as to have infinite values
even though they are finite. The simplest example is
\[\keyw{datatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
This is a finite codatatype, of cardinality one, whose single value is
infinite: $\mu\vthinspace a.\; A(a)$. Other specimens are
\[\begin{aligned}[t]
      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-\jot]
             \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
\end{aligned}
\]
where \ty{unit} is a datatype with the single constructor $\const{Unity} :
\ty{unit}$. We call such types \emph{corecursive singletons}. For the decision
procedure, it will be important to detect these types. A type may also be a
corecursive singleton only in some models. For example, if \ty{unit} is an
uninterpreted type, \ty{b} and \ty{c} will be singleton types precisely in the
models that interpret \ty{unit} as a singleton. Fortunately, given some
cardinalities for the ordinary types, it is easy to characterize this
degenerate case:

>>>>

Finally, some codatatypes are so degenerate as to have infinite values
even though they are finite. The simplest example is
\[\keyw{codatatype}\;\, \ty{a} \,=\, \const{A}(\ty{a})\]
This is a finite codatatype, of cardinality one, whose single value is
infinite: \,$\mu\vthinspace a.\; A(a)$. Other specimens are
\[\begin{aligned}[t]
      \keyw{codatatype}\;\, \ty{b} & \,=\, \const{B}(\ty{b},\: \ty{c},\: \ty{b},\: \ty{unit}) \\[-\jot]
             \keyw{and}\;\, \ty{c} & \,=\, \const{C}(\ty{a},\: \ty{unit},\: \ty{b},\: \ty{c})
\end{aligned}
\]
where \ty{unit} is a datatype with the single constructor $\const{Unity} :
\ty{unit}$. We call such types \emph{corecursive singletons}. For the decision
procedure, it will be important to detect these types. A type may also be a
corecursive singleton only in some models. For example, if \ty{unit} is an
uninterpreted type, \ty{b} and \ty{c} will be singleton types precisely in the
models that interpret \ty{unit} as a singleton. Fortunately, given some
cardinalities for the ordinary types, it is easy to characterize this
degenerate case:
